---
title: "Biobank"
author: "Alyssa Monda"
date: "9/21/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(knitr)
```

##Question: 
Does the relationship between depression and heartfailure differ based upon heart failure categorization. 

##Aims: 
1: Investigate the relationship between depression and future heart failure events. (Death, CV death, hospitalizations, Depression as measured by PHQ)

2:Investigate differences in the relationship between depression and future cardiac events varies based on heart failure categorization (ischemic vs. non ischemic HF, also HFpEF vs. HFrEF).

##Load data into R
```{r}
#Load data into R and assign data to the name CDT


```

##Filter Individuals from data set who meet inclusion criteria (Diagnosed with HF, Age 65+)
```{r}
#Subsset by age to ensure minimum age is 65
CD <-subset(CDT, age>=65)
summary(CD$age)

#Subset by HF (1=yes, 0=No)
CD <-subset(CDT, htfail=1)
summary(CD$htfail)

#Summarize by HF based on Chart Review Variable to make sure only those with dx of HF are included in the sample
C_htfail<- summary(as.factor(CD$htfail))
C_htfail
barplot(C_htfail)
```

*Include EF measures as inclusion criteria?*

##Data Cleaning and Initial Descriptive Statistics 
### Univariate Analysis Continuous Variables 
Distrbiution of Age
```{r}
summary(CD$age)
boxplot(CD$age)
hist(CD$age)
```

Distribution of Height (cm)
```{r}
summary(CD$subhtcm)
boxplot(CD$subhtcm)
hist(CD$subhtcm)
```

Distribution of Weight (kg)
```{r}
summary(CD$subwtkg)
boxplot(CD$subwtkg)
hist(CD$subwtkg)
```

Distribution of BMI
```{r}
summary(CD$blbmi)
boxplot(CD$blbmi)
hist(CD$blbmi)
```

Distribution of BSA
```{r}
summary(CD$bsa)
boxplot(CD$bsa)
hist(CD$bsa)
```

Distribution of Resting SBP
```{r}
summary(CD$restingsysbp)
boxplot(CD$restingsysbp)
hist(CD$restingsysbp)
```

Distribution of Resting DBP
```{r}
summary(CD$restingdiabp)
boxplot(CD$restingdiabp)
hist(CD$restingdiabp)
```

Distribution of Resting HR
```{r}
summary(CD$restinghr)
boxplot(CD$restinghr)
hist(CD$restinghr)
```

Distribution of Latest Ejection Fraction
```{r}
summary(CD$ef)
boxplot(CD$ef)
hist(CD$ef)
```

Distribution of Year of HF diagnosis
```{r}
summary(CD$hfdiag)
boxplot(CD$hfdiag)
hist(CD$hfdiag)
```

### Univariate Analysis Categorical Variables
Distribution of Gender
```{r}
#0=Female, 1=Male
C_gend<- summary(as.factor(CD$gend))
C_gend
barplot(C_gend)
```

Distribution of Hospital
```{r}
#0=EUH, 1=Midtown, 2=VA, 3=Grady, 4=Other
C_hosp<- summary(as.factor(CD$hosp))
C_hosp
barplot(C_hosp)
```

Distribution of Race
```{r}
#Race 1=causcasian, 2=AA, 3=Hispanic, 4=Asian, 5=Native American, 6=Pacific Islander, 7=Other
C_race<- summary(as.factor(CD$race))
C_race
barplot(C_race)

#Race_Other
C_raceO<- summary(as.factor(CD$raceoth))
C_raceO
barplot(C_raceO)
```

Distribution of Marital Status
```{r}
#1=married, 2=widowed, 3=divorced, 4=separated, 5=never married
C_ms<- summary(as.factor(CD$ms))
C_ms
barplot(C_ms)
```

Distribution of Employment Status
```{r}
#1=full time, 2=part time, 3=unemployed, 4=disabled, 5=retired
C_es<- summary(as.factor(CD$es))
C_es
barplot(C_es)
```

Distribution of Education
```{r}
#1=elementary/middle, 2=some high school, 3=HS grad, 4=some college, 5=college grad, 6=graduate education or degree
C_educ<- summary(as.factor(CD$educ))
C_educ
barplot(C_educ)
```

Distribution of Setting
```{r}
#1=inpt, 2=outpt
C_setting<- summary(as.factor(CD$setting))
C_setting
barplot(C_setting)
```

Distribution of Reason for Catheterization
```{r}
#Coded Reason
#1=positive stress test, 2=ST elevation MI, 3=Non-ST elevation myocardial, 4=unstable angina, 5=valuvular intervention, 6=peripheral intervention, 7=preop cardiac clearance, 8=evaluation for non-cardiac issues, 9=heart transplant, 10=HF, 11=Other, 12=TAVR
C_adm_reason <- summary(as.factor(CD$adm_reason))
C_adm_reason 
barplot(C_aadm_reason)

#Other Reason
C_admreason_other <- summary(as.factor(CD$admreason_other))
C_admreason_other 
barplot(C_admreason_other)
```

Distribution of Date of Catheterization
```{r}
#categorize by season?


#Summary
C_admcathdate<- summary(as.factor(CD$admcathdate))
C_admcathdate
barplot(C_admcathdate)
```

Distribution of Zip Code (How to analyze this? will it show up as numeric?)
```{r}
#Categorize into parts of atlanta?


#Summary
C_zip<- summary(as.factor(CD$zip))
C_zip
barplot(C_zip)
```

Distribution of NYHA
```{r}
#1=I can perform all physical activity without getting short of breath or tired, or having palpitations.
#2=I get short of breath or tired, or have palpitations when performing more strenuous activities. For example, walking on steep inclines or walking up several  ights of steps.
#3=I get short of breath or tired, or have palpitations when performing day to day activities. For example, walking on a  at surface.
#4=I feel breathless at rest, and am mostly housebound. I am unable to carry out any physical activity without getting short of breath or tired, or having palpitations.
C_sanyha<- summary(as.factor(CD$sanyha))
C_sanyha
barplot(C_sanyha)
```

Distribution of HF Type
```{r}
#1=systolic, 2=preserved
C_hftype<- summary(as.factor(CD$hftype))
C_hftype
barplot(C_hftype)
```

Distribution of HF Subtype
```{r}
#1=sIschemic Cardiomyopathy, 2=Non-ischemic, non-valvular Cardiomyopathy, 3=Valvular Disease-related Cardiomyopathy, 4=Genetic Syndrome, 5=Other
C_hftype<- summary(as.factor(CD$hftype))
C_hftype
barplot(C_hftype)
```

*Haven't set codes yet*

Distribution of Comorbidities (Choose some?)
```{r}
#
C_setting<- summary(as.factor(CD$setting))
C_setting
barplot(C_setting)
```

Distribution of Medications
```{r}
#1=inpt, 2=outpt
C_setting<- summary(as.factor(CD$setting))
C_setting
barplot(C_setting)
```


#Univariate Analysis of Outcome Variables
Distribution of Patient Status
```{r}
#0=alive, 1= dead, 2=unknown
C_pstatusyr<- summary(as.factor(CD$pstatusyr))
C_pstatusyr
barplot(C_pstatusyr)
```

Distribution of Source of Death Record
```{r}
#1=family contact, 2=chart review, 3=SSDI/Internet Search, 4=National Death Index
C_deathrecordyr<- summary(as.factor(CD$deathrecordyr))
C_deathrecordyr
barplot(C_deathrecordyr)
```

Distribution of Date of death or last known alive
```{r}
#text entered redo code
C_lastaliveyr<- summary(as.factor(CD$lastaliveyr))
C_lastaliveyr
barplot(C_lastaliveyr)
```

Distribution of Cause of Death
```{r}
#1= Immediate cardiac death; within one hour after onset of symptoms
#2=Sudden cardiac death; within 24 hours
#3=Cardiac death; not sudden or immediate
#4=Vascular; non-cardiac
#5=De nite non-cardiovascular death
#6= suicide
#7= unknown
#8=Probably cardiac death
#9=Probable non-cardiovascular death
C_causeofdeathyr<- summary(as.factor(CD$causeofdeathyr))
C_causeofdeathyr
barplot(C_causeofdeathyr)
```

Distribution of Associated events leading to patient's death
```{r}
#1= Acute myocardial infarction
#2=CHF
#3=Arrythmia
#4=Stroke
#5=Cancer
#6=Trauma
#7=During Surgery
#8= During Percutaneous Intervention
#9= During Catheterization
#10=Post-Surgery
#11= Post-PI
#12=CardiogenicShock
#13=Other/Unknown
C_setting<- summary(as.factor(CD$setting))
C_setting
barplot(C_setting)
```

Distribution of How many hospitalizations since Enrollment 
```{r}
#code as text so recode to show text? then categorize into groups or merge hospitalized yes no
C_setting<- summary(as.factor(CD$setting))
C_setting
barplot(C_setting)

#Hospitalized Year (1=yes, 0=no)
barplot(hospitalizedyr1)
barplot(hospitalizedyr2)
barplot(hospitalizedyr3)
barplot(hospitalizedyr4)
barplot(hospitalizedyr5)
barplot(hospitalizedyr6)
barplot(hospitalizedyr7)
barplot(hospitalizedyr8)
barplot(hospitalizedyr9)
barplot(hospitalizedyr10)


#Hospitalized visit type (1=outpatient, 2=inpatient)
barplot(visitypeyr1)
barplot(visitypeyr2)
barplot(visitypeyr3)
barplot(visitypeyr4)
barplot(hvisitypeyr5)
barplot(visitypeyr6)
barplot(visitypeyr7)
barplot(visitypeyr8)
barplot(visitypeyr9)
barplot(visitypeyr10)

#Hospitalized Reason
#1=chest pain/angina
#2=MI
#3=CHF
#4=Cardiac Cath
#5=PCI
#6=PTA
#7=CABG
#8=Valve Surgery
#9=HT
#10=Arrythmia
#11=Cardioversion
#12=Pacemaker
#13=Defibrillator
#14=Stress Test
#15=Vascular
#16=Stroke
#17=TIA
#18=Other
barplot(hospreasonyr1)
barplot(hospreasonyr2)
barplot(hospreasonyr3)
barplot(hospreasonyr4)
barplot(hospreasonyr5)
barplot(hospreasonyr6)
barplot(hospreasonyr7)
barplot(hospreasonyr8)
barplot(hospreasonyr9)
barplot(hospreasonyr10)
```

#Questionnaire Analysis 

PHQ
```{r}
#gather a sum and find out if its weighted or if there is a scoring algorithm to add 


```

BHQ

QDASI

STRESS


##Bivariate Analysis by subgroup of HF


##Bivariate Analysis by Depression 
Look for significant correlations and identify potential covariates that I need to control for. 
```{r}
#Scatterplot BMI*AGE to observe for any trends
plot(Meta$BMI_CORRECTED, Meta$AGE_YEARS)
```

##Bivariate Analysis with outcome variable



##Create Table One



##Aim 1 Outcomes by Depression
```{r}

```



##Aim 2 Outcomes by Depression based on Type and Subtype of HF
```{r}

```

