---
title: "Biobank"
author: "Alyssa Monda"
date: "9/21/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(knitr)
```

##Question: 
Does the relationship between depression and heartfailure differ based upon heart failure categorization. 

##Aims: 
1: Investigate the relationship between depression and future heart failure events. (Death, CV death, hospitalizations, Depression as measured by PHQ)

2:Investigate differences in the relationship between depression and future cardiac events varies based on heart failure categorization (ischemic vs. non ischemic HF, also HFpEF vs. HFrEF).

##Load data into R

```{r}
#Load data into R and assign data to the name CD


```

## Univariate Analysis Continuous Variables 
Distriution of Age
```{r}
summary(CD$age)
boxplot(CD$age)
hist(CD$age)
```

Distriution of BMI
```{r}
summary(CD$blbmi)
boxplot(CD$blbmi)
hist(CD$blbmi)
```

Distriution of BSA
```{r}
summary(CD$bsa)
boxplot(CD$bsa)
hist(CD$bsa)
```

Distriution of Resting SBP
```{r}
summary(CD$restingsysbp)
boxplot(CD$restingsysbp)
hist(CD$restingsysbp)
```

Distriution of Resting DBP
```{r}
summary(CD$restingdiabp)
boxplot(CD$restingdiabp)
hist(CD$restingdiabp)
```

Distriution of Resting HR
```{r}
summary(CD$restinghr)
boxplot(CD$restinghr)
hist(CD$restinghr)
```


## Univariate Analysis Categorical Variables 
Distribution of Hospital
```{r}
C_hosp<- summary(as.factor(CD$hosp))
C_hosp
barplot(C_hosp)
```

Distribution of Zip Code
```{r}
C_zip<- summary(as.factor(CD$zip))
C_zip
barplot(C_zip)
```

Distribution of Race
```{r}
#Race
C_race<- summary(as.factor(CD$race))
C_race
barplot(C_race)

#Race_Other
C_raceO<- summary(as.factor(CD$raceoth))
C_raceO
barplot(C_raceO)
```


##Bivariate Analysis

```{r}
#Scatterplot BMI*AGE to observe for any trends
plot(Meta$BMI_CORRECTED, Meta$AGE_YEARS)
```

