---
title: "Final Abstract"
author: "Alyssa Monda"
date: "11/26/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(knitr)
library(psych)
```


##Load data into R
```{r}
#Load data into R and assign data to the name BB
library(readr)
BB <- read_csv("~/Desktop/Biobank.csv", 
    col_types = cols(hfage = col_double()))
View(BB)
```

##Filter Individuals from data set who meet inclusion criteria (Diagnosed with HF, "age" 65+)
```{r}
#Subsset by "age" to ensure minimum "age" is 65
OBB <-subset(BB, age>=65)
summary(OBB$`age`)

#Subset by HF (1=yes, 0=No) ****Total N for OA with HF = 478 *****
CVBB <-subset(OBB, hfhist== 1)
summary(CVBB$hfhist)
CVBB$hfhist

```

#Distribution of depression diagnosis vs depressive symtpoms 
```{r}
hist(CVBB$dephist)
hist(CVBB$`mdclass#`)
#Based on distribution it appears it may be best to use depressive symptoms and create 3 groups, minimal to mild, moderate,and  moderately severe to severe. Then compare outcomes of the 3 groups. There are a limited number with diagnosed depression which does not match PHQ results. 

#Created subgroup column of groups in excel. Import data back into R.
#Import CVBB
library(readr)
CVBB <- read_csv("~/Desktop/AmericanGut/Big Data/Health_Outcomes/CVBB.csv")
View(CVBB)

#Histogram showing groups of min to mild, mild to moderate, moderate to severe
hist(CVBB$mdgroup)

#Subset minimal to mild depression
DMin <-subset(CVBB,mdtype<= 2)
summary(DMin$'mdclass#')
DMin$'mdclass#'

#Subset moderate depression
DMod <-subset(CVBB,mdtype== 3)
summary(DMod$'mdclass#')
DMod$'mdclass#'

#Subset moderately severe to severe depression
DSev <-subset(CVBB,mdtype>=4)
summary(DSev$'mdclass#')
DSev$'mdclass#'


```

Total N=478
Min-Mild Depressive Sx n=124
Moderate Depressive Sx n=160
Moderately Severe to Severe Depressive Sx n=131

#Summary of total dataset
```{r}
summary(CVBB,  na.rm=TRUE)
describe(CVBB, na.rm=TRUE)
```

#Summary of Min to Mild
```{r}
summary(DMin,  na.rm=TRUE)
describe(DMin,  na.rm=TRUE)
```

#Summary of Moderate
```{r}
summary(DMod,  na.rm=TRUE)
describe(DMod,  na.rm=TRUE)
```

#Summary of Moderately Severe to Severe 
```{r}
summary(DSev,  na.rm=TRUE)
describe(DSev,  na.rm=TRUE)
```


#Turn Numeric variables into factor
```{r}
CVBB <- within(CVBB, {
  aahist <- as.factor(aahist)
  adace <- as.factor(adace)
  adarb <- as.factor(adarb)
  adasp <- as.factor(adasp)
  adbb <- as.factor(adbb)
  adcaantag <- as.factor(adcaantag)
  adcou <- as.factor(adcou)
  adeze <- as.factor(adeze)
  adfib <- as.factor(adfib)
  adfuro <- as.factor(adfuro)
  adhctz <- as.factor(adhctz)
  adm_HF <- as.factor(adm_HF)
  adm_NonCardiac <- as.factor(adm_NonCardiac)
  adm_NSTEMI <- as.factor(adm_NSTEMI)
  adm_Other <- as.factor(adm_Other)
  adm_PerInterv <- as.factor(adm_PerInterv)
  adm_PosStressTest <- as.factor(adm_PosStressTest)
  adm_PreOpClear <- as.factor(adm_PreOpClear)
  adm_STEMI <- as.factor(adm_STEMI)
  adm_TAVR <- as.factor(adm_TAVR)
  adm_Transplant <- as.factor(adm_Transplant)
  adm_UnAngina <- as.factor(adm_UnAngina)
  adm_Valvular <- as.factor(adm_Valvular)
  admet <- as.factor(admet)
  adniacin <- as.factor(adniacin)
  adnitrate <- as.factor(adnitrate)
  adpplavix <- as.factor(adpplavix)
  adspir <- as.factor(adspir)
  adstat <- as.factor(adstat)
})
```

```{r}
CVBB <- within(CVBB, {
  apnhist <- as.factor(apnhist)
  arrhist <- as.factor(arrhist)
  ashist <- as.factor(ashist)
  behq2 <- as.factor(behq2)
  behq6 <- as.factor(behq6)
  behq8 <- as.factor(behq8)
  behq9 <- as.factor(behq9)
  behq10 <- as.factor(behq10)
  behq14 <- as.factor(behq14)
})
```

```{r}
CVBB <- within(CVBB, {
  cabg <- as.factor(cabg)
  cahist <- as.factor(cahist)
  canhist <- as.factor(canhist)
})
```

```{r}
CVBB <- within(CVBB, {
  CHDdeath <- as.factor(CHDdeath)
  CHDdeath1yr <- as.factor(CHDdeath1yr)
  CHDdeath3yr <- as.factor(CHDdeath3yr)
  CHDdeath5yr <- as.factor(CHDdeath5yr)
  CHF <- as.factor(CHF)
  copdhist <- as.factor(copdhist)
  cphist <- as.factor(cphist)
  CVdeath <- as.factor(CVdeath)
  CVDeath1yr <- as.factor(CVDeath1yr)
  CVDeath3yr <- as.factor(CVDeath3yr)
  CVDeath5yr <- as.factor(CVDeath5yr)
  death <- as.factor(death)
  death1yr <- as.factor(death1yr)
  death3yr <- as.factor(death3yr)
  death5yr <- as.factor(death5yr)
  dephist <- as.factor(dephist)
  dmhist <- as.factor(dmhist)
  educ <- as.factor(educ)
  es <- as.factor(es)
  hfage <- as.factor(hfage)
  hfhist <- as.factor(hfhist)
  histang <- as.factor(histang)
  histcath <- as.factor(histcath)
  histstent <- as.factor(histstent)
  hldhist <- as.factor(hldhist)
  hosp <- as.factor(hosp)
  hospcard <- as.factor(hospcard)
  hospncard <- as.factor(hospncard)
  htnhist <- as.factor(htnhist)
  htvalv <- as.factor(htvalv)
  ibdhist <- as.factor(ibdhist)
  icd <- as.factor(icd)
  icshist <- as.factor(icshist)
  incp <- as.factor(incp)
  inpalp <- as.factor(inpalp)
  insob <- as.factor(insob)
  insyncope <- as.factor(insyncope)
  lhc_adm <- as.factor(lhc_adm)
  MACE <- as.factor(MACE)
  MACE1yr <- as.factor(MACE1yr)
  MACE3yr <- as.factor(MACE3yr)
  MACE5yr <- as.factor(MACE5yr)
  MALE_Comb <- as.factor(MALE_Comb)
  MALECV <- as.factor(MALECV)
  MALECV_Comb <- as.factor(MALECV_Comb)
  mdappt <- as.factor(mdappt)
  mdbad <- as.factor(mdbad)
  mdclass. <- as.factor(mdclass.)
  mdconc <- as.factor(mdconc)
  mdcouns <- as.factor(mdcouns)
  mddead <- as.factor(mddead)
  mddep <- as.factor(mddep)
  mdplea <- as.factor(mdplea)
  mdsleep <- as.factor(mdsleep)
  mdspeak <- as.factor(mdspeak)
  mdtired <- as.factor(mdtired)
})
```

```{r}
CVBB <- within(CVBB, {
  MI <- as.factor(MI)
  mihist <- as.factor(mihist)
  ms <- as.factor(ms)
  nonCVdeath <- as.factor(nonCVdeath)
  padhist <- as.factor(padhist)
  pci_adm <- as.factor(pci_adm)
  permpace <- as.factor(permpace)
  PTA <- as.factor(PTA)
  PTA_Comb <- as.factor(PTA_Comb)
  pvd_adm <- as.factor(pvd_adm)
  PVDCPT <- as.factor(PVDCPT)
  race <- as.factor(race)
})
```

```{r}
CVBB <- within(CVBB, {
  REVASC <- as.factor(REVASC)
  rhc_adm <- as.factor(rhc_adm)
  sanyha <- as.factor(sanyha)
  saqnitro <- as.factor(saqnitro)
  saqtime <- as.factor(saqtime)
  Sertraline <- as.factor(Sertraline)
  setting <- as.factor(setting)
  slehist <- as.factor(slehist)
  smoking <- as.factor(smoking)
  strhist <- as.factor(strhist)
  STROKE <- as.factor(STROKE)
})
```

```{r}
CVBB <- within(CVBB, {
  thyhist <- as.factor(thyhist)
})
```

```{r}
CVBB <- within(CVBB, {
  transp <- as.factor(transp)
})
```

```{r}
CVBB <- within(CVBB, {
  valhist <- as.factor(valhist)
  valv_adm <- as.factor(valv_adm)
  vtehist <- as.factor(vtehist)
})
```


#ANOVA Tests between groups and numeric variables
```{r}
AnovaModel.3 <- aov(age ~ mdgroup, data=CVBB)
summary(AnovaModel.3)
with(CVBB, numSummary(age, groups=mdgroup, statistics=c("mean", "sd")))
```

```{r}
AnovaModel.4 <- aov(blbmi ~ mdgroup, data=CVBB)
summary(AnovaModel.4)
with(CVBB, numSummary(blbmi, groups=mdgroup, statistics=c("mean", "sd")))
```

```{r}
AnovaModel.5 <- aov(bsa ~ mdgroup, data=CVBB)
summary(AnovaModel.5)
with(CVBB, numSummary(bsa, groups=mdgroup, statistics=c("mean", "sd")))
```

```{r}
AnovaModel.6 <- aov(restingdiabp ~ mdgroup, data=CVBB)
summary(AnovaModel.6)
with(CVBB, numSummary(restingdiabp, groups=mdgroup, statistics=c("mean", 
  "sd")))
```

```{r}
AnovaModel.7 <- aov(restinghr ~ mdgroup, data=CVBB)
summary(AnovaModel.7)
with(CVBB, numSummary(restinghr, groups=mdgroup, statistics=c("mean", 
  "sd")))
```

```{r}
AnovaModel.8 <- aov(restingsysbp ~ mdgroup, data=CVBB)
summary(AnovaModel.8)
with(CVBB, numSummary(restingsysbp, groups=mdgroup, statistics=c("mean", 
  "sd")))
```

```{r}
AnovaModel.9 <- aov(timetodeath ~ mdgroup, data=CVBB)
summary(AnovaModel.9)
with(CVBB, numSummary(timetodeath, groups=mdgroup, statistics=c("mean", 
  "sd")))
```

```{r}
AnovaModel.10 <- aov(timetodeath1yr ~ mdgroup, data=CVBB)
summary(AnovaModel.10)
with(CVBB, numSummary(timetodeath1yr, groups=mdgroup, statistics=c("mean", 
  "sd")))
```

```{r}
AnovaModel.11 <- aov(timetodeath3yr ~ mdgroup, data=CVBB)
summary(AnovaModel.11)
with(CVBB, numSummary(timetodeath3yr, groups=mdgroup, statistics=c("mean", 
  "sd")))
```

```{r}
AnovaModel.12 <- aov(timetodeath5yr ~ mdgroup, data=CVBB)
summary(AnovaModel.12)
with(CVBB, numSummary(timetodeath5yr, groups=mdgroup, statistics=c("mean", 
  "sd")))
```

```{r}
AnovaModel.13 <- aov(timetohospcard ~ mdgroup, data=CVBB)
summary(AnovaModel.13)
with(CVBB, numSummary(timetohospcard, groups=mdgroup, statistics=c("mean", 
  "sd")))
```

```{r}
AnovaModel.14 <- aov(timetohospncard ~ mdgroup, data=CVBB)
summary(AnovaModel.14)
with(CVBB, numSummary(timetohospncard, groups=mdgroup, statistics=c("mean", 
  "sd")))
```

#ChiSq tests between groups and categorical variables
```{r}
local({
  .Table <- xtabs(~mdgroup+race, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```

```{r}
local({
  .Table <- xtabs(~mdgroup+ms, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```

```{r}
local({
  .Table <- xtabs(~mdgroup+es, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```

```{r}
local({
  .Table <- xtabs(~mdgroup+educ, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```

```{r}
local({
  .Table <- xtabs(~mdgroup+sanyha, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+smoking, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+htnhist, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+dmhist, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```

```{r}
local({
  .Table <- xtabs(~mdgroup+copdhist, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```

```{r}
local({
  .Table <- xtabs(~mdgroup+canhist, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+cahist, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+strhist, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+dephist, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+apnhist, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```

```{r}
local({
  .Table <- xtabs(~mdgroup+ibdhist, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+thyhist, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+hldhist, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+mdcouns, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+cabg, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+histcath, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+histang, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+histstent, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+transp, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+htvalv, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+permpace, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+icd, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+adace, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+adhctz, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+adfuro, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+adspir, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+adbb, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+adcaantag, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+adasp, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```

```{r}
local({
  .Table <- xtabs(~mdgroup+adcou, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```

```{r}
local({
  .Table <- xtabs(~mdgroup+adpplavix, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+adnitrate, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+admet, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+causeofdeathyr, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+hospcard, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+hospncard, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+CVdeath, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
```{r}
local({
  .Table <- xtabs(~mdgroup+nonCVdeath, data=CVBB)
  cat("\nFrequency table:\n")
  print(.Table)
  cat("\nRow percentages:\n")
  print(rowPercents(.Table))
  .Test <- chisq.test(.Table, correct=FALSE)
  print(.Test)
})
```
##Univariate
Skewness and Kurtosis observed in the describe command
```{r}
#could also view histograms
ggplot(CVBB, aes(x=age)) + 
  geom_histogram(aes(y=..density..), 
                 binwidth=5,
                 colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666") +
  labs(title="Histogram of Age")

ggplot(CVBB, aes(x=blbmi)) + 
  geom_histogram(aes(y=..density..), 
                 binwidth=5,
                 colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666") +
  labs(title="Histogram of BMI")

#Significant difference between group but ~normally distributed and w/in normal clinical range
ggplot(CVBB, aes(x=restinghr)) + 
  geom_histogram(aes(y=..density..), 
                 binwidth=5,
                 colour="black", fill="white") +
  geom_density(alpha=.2, fill="#FF6666") +
  labs(title="Histogram of HR")

```

###Combine marital Status variables
Partnered vs non-partnered 
```{r}
CVBB$Single <- CVBB$ms
CVBB$Single

dataA$var1r <- factor(dataA$var1r,
                      levels=c(0,1),
                      labels=c("NO var1 <= 3",
                               "YES var1 > 3")) 
```

Now check for significant difference between groups again using chisq 
```{r}
# the correlations
psych::corr.test(helpset1[,vars],
                 method="pearson")

# save the results
pc <- psych::corr.test(helpset1[,vars],
                       method="pearson")
# look at correlations
pc$r
```



## Bivariate Stats
Continuous using Pearson R
```{r}
# the correlations
psych::corr.test(CVBB[,vars],
                 method="pearson")

# save the results
pc <- psych::corr.test(CVBB[,vars],
                       method="pearson")
# look at correlations
pc$r

# t-test for each correlation
pc$t

# p-value of each t-test for correlation
pc$p

# confidence intervals for each pair
# removes the diagonal values and tests
pc$ci

# also look at spearmans rho
pc <- psych::corr.test(helpset1[,vars],
                       method="spearman")
pc$ci

```


Categorical using t test or ANOVA




















